---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit the README.Rmd file -->


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Lab report #3 - instructions

Follow the instructions posted at https://ds202-at-isu.github.io/labs.html for the lab assignment. The work is meant to be finished during the lab time, but you have time until Monday evening to polish things.

Include your answers in this document (Rmd file). Make sure that it knits properly (into the md file). Upload both the Rmd and the md file to your repository.  

All submissions to the github repo will be automatically uploaded for grading once the due date is passed.
Submit a link to your repository on Canvas (only one submission per team) to signal to the instructors that you are done with your submission.

# Lab 3: Avenger's Peril

## As a team

Extract from the data below two data sets in long form `deaths` and `returns`

```{r}
av <- read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/avengers/avengers.csv", stringsAsFactors = FALSE)
head(av)
```

Get the data into a format where the five columns for Death[1-5] are replaced by two columns: Time, and Death. Time should be a number between 1 and 5 (look into the function `parse_number`); Death is a categorical variables with values "yes", "no" and "". Call the resulting data set `deaths`. 

```{r}
library(dplyr)
library(tidyverse)

av_deaths <- av %>%
  gather(key = "Time", value = "Death", starts_with("Death")) %>%
  mutate(Time = as.integer(gsub("Death", "", Time)), 
         Death = case_when(
           Death == "YES" ~ "yes",
           Death == "NO" ~ "no",
           TRUE ~ ""
         ))

deaths_per_avenger <- av_deaths %>%
  filter(Death == "yes") %>%
  group_by(Name.Alias) %>%
  summarize(Deaths = n())

average_deaths <- mean(deaths_per_avenger$Deaths)
average_deaths


```


Similarly, deal with the returns of characters.

Based on these datasets calculate the average number of deaths an Avenger suffers. 

## Individually

For each team member, copy this part of the report. 

Each team member picks one of the statements in the FiveThirtyEight [analysis](https://fivethirtyeight.com/features/avengers-death-comics-age-of-ultron/) and fact checks it based on the data. Use dplyr functionality whenever possible.

### FiveThirtyEight Statement

> Quote the statement you are planning to fact-check.
"Given the Avengers’ 53 years in operation and overall mortality rate, fans of the comics can expect one current or former member to die every seven months or so, with a permanent death occurring once every 20 months."


### Include the code (Nicholas Olech)
```{r}

#Calculate the total amount of deaths
total_deaths <- sum(av$Death1 =="YES"| av$Death2 =="YES" | av$Death3 =="Yes"|av$Death4 =="Yes"|av$Death5 =="YES")
#total_deaths

#Calculate the total amount of returns
total_returns <- sum(av$Return1 =="YES"| av$Return2 =="YES" | av$Return3 =="Yes"|av$Return4 =="Yes"|av$Return5 =="YES")
#total_returns

#Find the total amount of permanent deaths
perm_deaths <- total_deaths - total_returns

#Avengers have existed for 53 years, convert that to months to verify statistic
months_of_avengers <- 53 * 12

#To see how often an avenger is  killed, but not permanently
killed_frequency <- months_of_avengers / total_deaths
killed_frequency

#To see how often an avenger is killed permanently
perm_killed_frequency <- months_of_avengers / perm_deaths
perm_killed_frequency


```
Make sure to include the code to derive the (numeric) fact for the statement

### Include your answer (Nicholas Olech)

Based on my analysis, this statistic appears to be misleading, likely due to a discrepancy between the date the article was published and when the data set was updated. Instead of a death occurring 7 months, like the article stated, it appears to happen just once every ~9 months. Also the permanent death occurs only once every ~28 months which is also different from the articles statistic of 20.


### FiveThirtyEight Statement (Carter Parks)

> There’s a 2-in-3 chance that a member of the Avengers returned from their first stint in the afterlife, but only a 50 percent chance they recovered from a second or third death.8

### Include the code (Carter Parks)

```{r}
death1Return <- av %>%
  filter(Death1 == 'YES', Return1 == 'YES') %>%
  nrow()

death1Dead <- av %>%
  filter(Death1 == 'YES', Return1 == 'NO') %>%
  nrow()

death1Total <- death1Return + death1Dead

death2Return <- av %>%
  filter(Death2 == 'YES', Return2 == 'YES') %>%
  nrow()

death2Dead <- av %>%
  filter(Death2 == 'YES', Return2 == 'NO') %>%
  nrow()

death2Total <- death2Return + death2Dead

death3Return <- av %>%
  filter(Death2 == 'YES', Return2 == 'YES') %>%
  nrow()

death3Dead <- av %>%
  filter(Death2 == 'YES', Return2 == 'NO') %>%
  nrow()

death3Total <- death3Return + death3Dead

print(paste("Percentage returning from death 1:", death1Return / death1Total))
print(paste("Percentage returning from death 2:", death2Return / death2Total))
print(paste("Percentage returning from death 3:", death3Return / death3Total))

```
Make sure to include the code to derive the (numeric) fact for the statement

### Include your answer (Carter Parks)
Exactly 2/3s of avengers came back from their first death, while only 50% of avengers came back from their 2nd and 3rd deaths.

Upload your changes to the repository. Discuss and refine answers as a team.

### FiveThirtyEight Statement (Chris Draper)

> "Out of 173 listed Avengers, my analysis found that 69 had died at least one time after they joined the team.5 That’s about 40 percent of all people who have ever signed on to the team." 

### Include the code (Chris Draper)

```{r}

avengers_deaths <- av %>%
  
  mutate(Has_Died = if_else(Death1 == "YES" | Death2 == "YES" | Death3 == "YES" | Death4 == "YES" | Death5 == "YES", TRUE, FALSE)) %>%
 
  distinct(Name.Alias, .keep_all = TRUE)


total_avengers = nrow(avengers_deaths)

avengers_died_once = sum(avengers_deaths$Has_Died)

percentage_died_once = (avengers_died_once / total_avengers) * 100


cat("Out of", total_avengers, "listed Avengers,", avengers_died_once, "have died at least once after they joined the team.\n")

cat("That’s about", round(percentage_died_once, 2), "percent of all people who have ever signed on to the team.")

```
Make sure to include the code to derive the (numeric) fact for the statement

### Include your answer (Chris Draper)

- I am calculating less deaths and slightly less death rate than 40%. However, the original statement is true enough and not misleading. 

Upload your changes to the repository. Discuss and refine answers as a team.
